Estimating demographic bias on tests of children's early vocabulary

### 2022-01-21 paper/cogsci.Rmd
added pruning vs. demo effect size plot

mm_sex <- extract_group_df(mod_intuitive_sex, groups=c("Male","Female"))
mm_ses <- extract_group_df(mod_intuitive_ses, groups=c("low","high"))
mm_eth <- extract_group_df(mod_intuitive_eth, groups=c("Nonwhite","White"))

# varying thresholds
prune_dat <- tibble()
for(thresh in seq(.25, 2.5, .25)) {
  tmp_sex <- get_extreme_item_difficulty_differences(mm_sex, sd_thresh=thresh) # min: -.53 max: 1.06
  tmp_ses <- get_extreme_item_difficulty_differences(mm_ses, sd_thresh=thresh) # min: -1.07 max: 1.13
  tmp_eth <- get_extreme_item_difficulty_differences(mm_eth, sd_thresh=thresh) # min: -.64 max: 1.56
  prune_defs <- intersect(intersect(subset(tmp_ses, extreme)$definition, 
                       subset(tmp_eth, extreme)$definition),
                       subset(tmp_sex, extreme)$definition)

@@

# 2SD threshold
mm_sex2 <- get_extreme_item_difficulty_differences(mm_sex) # min: -.53 max: 1.06
mm_ses2 <- get_extreme_item_difficulty_differences(mm_ses) # min: -1.07 max: 1.13
mm_eth2 <- get_extreme_item_difficulty_differences(mm_eth) # min: -.64 max: 1.56

all_extreme <- rbind(subset(mm_sex2, extreme),
                     subset(mm_ses2, extreme),
                     subset(mm_eth2, extreme)) 
# length(unique(all_extreme$definition)) 
# sort(table(all_extreme$definition)) # candy, choo choo, daddy, duck, ... tractor, vroom

@@

>>>>

mm_sex <- extract_group_df(mod_intuitive_sex, groups=c("Male","Female"))
mm_ses <- extract_group_df(mod_intuitive_ses, groups=c("low","high"))
mm_eth <- extract_group_df(mod_intuitive_eth, groups=c("Nonwhite","White"))
```

```{r train-pruned-models, eval=F}
# varying thresholds
prune_dat <- tibble()
for(thresh in seq(.25, 3.0, .25)) {
  tmp_sex <- get_extreme_item_difficulty_differences(mm_sex, sd_thresh=thresh) # min: -.53 max: 1.06
  tmp_ses <- get_extreme_item_difficulty_differences(mm_ses, sd_thresh=thresh) # min: -1.07 max: 1.13
  tmp_eth <- get_extreme_item_difficulty_differences(mm_eth, sd_thresh=thresh) # min: -.64 max: 1.56
  prune_defs <- unique(c(subset(tmp_ses, extreme)$definition, 
                       subset(tmp_eth, extreme)$definition,
                       subset(tmp_sex, extreme)$definition))

@@

save(prune_dat, file=here("data/pruned_demo_coefs.Rdata"))
```

```{r}
# 2SD threshold
mm_sex <- get_extreme_item_difficulty_differences(mm_sex) # min: -.53 max: 1.06
mm_ses <- get_extreme_item_difficulty_differences(mm_ses) # min: -1.07 max: 1.13
mm_eth <- get_extreme_item_difficulty_differences(mm_eth) # min: -.64 max: 1.56

all_extreme <- rbind(subset(mm_sex, extreme),
                     subset(mm_ses, extreme),
                     subset(mm_eth, extreme)) 
# length(unique(all_extreme$definition)) 
# sort(table(all_extreme$definition)) # candy, choo choo, daddy, duck, ... tractor, vroom

@@

```{r pruned-coefs, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.5, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Size of demographic effects (regression coefficients) vs. the threshold of pruned items."}
load(here("data/pruned_demo_coefs.Rdata"))

prune_dat %>% rename(c(beta_sex=sex_beta, beta_ses=ses_beta, beta_eth=eth_beta)) %>%
  pivot_longer(c("extreme_sex","extreme_ses","extreme_eth","beta_sex","beta_ses","beta_eth"), 
                           names_to=c("type", "demo"), names_sep="_") %>%
  pivot_wider(names_from=type, values_from=value) %>%
  mutate(Dimension = ifelse(demo=="ses", "maternal ed", ifelse(demo=="eth", "race", demo))) %>%
  ggplot(aes(x=SDthreshold, y=beta, color=extreme)) + facet_wrap(. ~ Dimension) + 
  geom_point() + theme(legend.position = "bottom") +
  labs(color="Excluded items") + 
  xlab("Exclusion Threshold (SDs)") + ylab("Regression coefficient (beta)")
```
